---
title: 'Statistical Inference Course Project'
date: '19/09/2021'
author: 'Yulong Wang'
output:
  pdf_document: 
    latex_engine: xelatex
  html_document:
    keep_md: yes
---

```{r Folder}
setwd("/Users/yulong/GitHub/Statistical-Inference-Course-Project")
```

# Part 1: Simulation Exercise Instructions

## Instructions

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should

* Show the sample mean and compare it to the theoretical mean of the distribution.

* Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

* Show that the distribution is approximately normal.

## Loading Libraries

```{r DataLoading, results="hide"}
library("data.table")
library("ggplot2")
```

## Task

```{r Stuff}
# set seed for reproducability
set.seed(31)
# set lambda to 0.2
lambda <- 0.2
# 40 samples
n <- 40
# 1000 simulations
simulations <- 1000
# simulate
simulated_exponentials <- replicate(simulations, rexp(n, lambda))
# calculate mean of exponentials
means_exponentials <- apply(simulated_exponentials, 2, mean)
```

## Question 1
Show where the distribution is centered at and compare it to the theoretical center of the distribution.

```{r}
analytical_mean <- mean(means_exponentials)
analytical_mean
```

```{r}
# analytical mean
theory_mean <- 1/lambda
theory_mean
```

```{r}
# visualization
hist(means_exponentials, xlab = "mean", main = "Exponential Function Simulations")
abline(v = analytical_mean, col = "red")
abline(v = theory_mean, col = "orange")
```

The analytics mean is 4.993867 the theoretical mean 5. The center of distribution of averages of 40 exponentials is very close to the theoretical center of the distribution.

## Question 2

Show how variable it is and compare it to the theoretical variance of the distribution.. 

```{r}
# standard deviation of distribution
standard_deviation_dist <- sd(means_exponentials)
standard_deviation_dist
```

```{r}
# standard deviation from analytical expression
standard_deviation_theory <- (1/lambda)/sqrt(n)
standard_deviation_theory
```

```{r}
# variance of distribution
variance_dist <- standard_deviation_dist^2
variance_dist
```

```{r}
# variance from analytical expression
variance_theory <- ((1/lambda)*(1/sqrt(n)))^2
variance_theory
```

Standard Deviation of the distribution is 0.7931608 with the theoretical SD calculated as 0.7905694. The Theoretical variance is calculated as ((1 / lambda) * (1/n2))2 = 0.625. The actual variance of the distribution is 0.6291041

## Question 3
Show that the distribution is approximately normal.

```{r}
xfit <- seq(min(means_exponentials), max(means_exponentials), length=100)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(n)))
hist(means_exponentials,breaks=n,prob=T,col="orange",
     xlab = "means",main="Density of means",ylab="density")
lines(xfit, yfit, pch=22, col="black", lty=5)
```

```{r}
# compare the distribution of averages of 40 exponentials to a normal distribution
qqnorm(means_exponentials)
qqline(means_exponentials, col = 2)
```

Due to Due to the central limit theorem (CLT), the distribution of averages of 40 exponentials is very close to a normal distribution.